---
title: "Predicting Churn using Social Network Analysis with R"
author: "Lenox Low"
date: "September 28, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This tutorial used the sample data to demonstrate the basic processes of performing social network analysis with R. The `igraph` library was used to do the social network analysis. You should install the latest version of `igraph` for R.
```{r eval=TRUE, echo=FALSE}
```

##(A) Import Libraries
```{r eval=TRUE, echo=TRUE, message= FALSE}
library("igraph")
library("data.table")
```


##(B) Import Data
```{r eval=TRUE, echo = TRUE}
transaction <- read.csv("../Data/transaction_sample.csv", header=TRUE);
head(transaction)
```


##(C) Data Preprocessing
Extract the unique MSISDN from the transaction data, then calculate the total incoming and outgoing call count and duration for each MSISDN. Two functions were created to extract the MSISDNs.

###(I) Extract unique MSISDN from the transaction data
```{r eval=TRUE, echo=TRUE}

ExtractUniqueMSISDNList <- function(target.df) {
  temp1 <- unique(target.df$PARTY_A)
  temp2 <- unique(target.df$PARTY_B)
  temp <- c(temp1, temp2)
  
  m.list <- unique(temp)
  m.list <- data.frame(m.list)
  colnames(m.list)[1] <- "MSISDN"
  
  return(m.list)
}
subs <- ExtractUniqueMSISDNList(transaction)
```

###(II) Extract the transaction reecords of defined MSISDN.
```{r eval=TRUE, echo=TRUE}
ExtractTransactionByMSISDN <- function(msisdn, target.df){
  sub.trans <- target.df[target.df$PARTY_A == msisdn | target.df$PARTY_B == msisdn, ]
  
  return(sub.trans)
}
```

Calculate the total call count and duration
```{r eval=TRUE, echo=TRUE}


for(i in 1:nrow(subs)){
  subs$CALL_IN_DUR[i] <- sum(transaction$CALL_DURATION[transaction$PARTY_B == subs$MSISDN[i]]) 
  subs$CALL_OUT_DUR[i] <- sum(transaction$CALL_DURATION[transaction$PARTY_A == subs$MSISDN[i]]) 
  
  subs$CALL_IN_COUNT[i] <- sum(transaction$CALL_COUNT[transaction$PARTY_B == subs$MSISDN[i]]) 
  subs$CALL_OUT_COUNT[i] <- sum(transaction$CALL_COUNT[transaction$PARTY_A == subs$MSISDN[i]]) 
  
  subs$RELATION_COUNT[i] <- nrow(ExtractUniqueMSISDNList(ExtractTransactionByMSISDN(subs$MSISDN[i], transaction))) - 1 
  subs$IN_RELATION[i] <- nrow(transaction[transaction$PARTY_B == subs$MSISDN[i], ])
  subs$OUT_RELATION[i] <- nrow(transaction[transaction$PARTY_A == subs$MSISDN[i], ])
}

subs$TOTAL_CALL_COUNT <- subs$CALL_IN_COUNT + subs$CALL_OUT_COUNT

head(subs)
```

###(I) Data Cleansing
Set the boundary value for the cleansing criteria
```{r eval=TRUE, echo=TRUE}
call.relation <- 30
call.count <- 30
call.mean <- mean(subs$TOTAL_CALL_COUNT)
call.duration <- 15
```

Label every MSIDSN based on the criteria stated below.
CALL CENTER - Incoming call is greater than call count boundary
              No outgoing call
              Call relation count is greater than call relation count boundary
TELESALES -   Outgoing call is greater than call count boundary
              No incoming call
              Outgoing call divided by call relation count is less than the mean of the total call count
NOISY -       Incoming call count is 1
              No outgoing call
              The call duration is less than the call duration boundary
```{r eval=TRUE, echo=TRUE}
for(i in 1: nrow(subs)){
  if(subs$CALL_IN_COUNT[i] > call.count && subs$CALL_OUT_COUNT[i] == 0 && subs$RELATION_COUNT[i] > call.relation) {
    subs$TYPE[i] <- "CALL CENTER"
  } else if (subs$CALL_IN_COUNT[i] == 0 && subs$CALL_OUT_COUNT[i] > call.count && subs$CALL_OUT_COUNT[i] / subs$RELATION_COUNT[i] < call.mean) {
    subs$TYPE[i] <- "TELESALES"
  } else if (subs$CALL_IN_COUNT[i] == 1 && subs$CALL_OUT_COUNT[i] == 0 && subs$CALL_IN_DUR[i] < call.duration) {
    subs$TYPE[i] <- "NOISY"
  } else {
    subs$TYPE[i] <- "NORMAL"
  }
}

normal.msisdn <- subs[subs$TYPE == "NORMAL",]
```

Get the transaction records of normal MSISDN only and calculate average duration per call (Weight) betwen two MSISDNs.
```{r eval=TRUE, echo=TRUE}
filtered.transaction <- transaction[transaction$PARTY_A %in% normal.msisdn$MSISDN & transaction$PARTY_B %in% normal.msisdn$MSISDN,]
filtered.transaction$Weight <- filtered.transaction$CALL_DURATION / filtered.transaction$CALL_COUNT
head(filtered.transaction)
```

###(II) Transform Data Frame to Matrix
```{r eval=TRUE, echo=TRUE}
msisdn.list <- ExtractUniqueMSISDNList(filtered.transaction)
ma <- matrix(0, ncol=nrow(msisdn.list), nrow=nrow(msisdn.list))
madf <- data.frame(ma)

colnames(madf) <- msisdn.list$MSISDN
row.names(madf) <- msisdn.list$MSISDN

filtered.transaction$PARTY_A <- as.character(filtered.transaction$PARTY_A)
filtered.transaction$PARTY_B <- as.character(filtered.transaction$PARTY_B)

for(i in 1:nrow(filtered.transaction)) {
  madf[filtered.transaction[i,"PARTY_A"], filtered.transaction[i,"PARTY_B"]] <- filtered.transaction[i, "Weight"]
}
ma <- as.matrix(madf)
ma[1:6, 1:6]
```


##(D) Social Network Analysis (SNA)
###(I) Build the network
```{r eval=TRUE, echo=TRUE, message=FALSE}
net <- graph.adjacency(ma, mode="directed", weighted=TRUE, diag=FALSE)
```

###(II) Generate Result
####(1) Closeness Centrality - how fast can a node reach everyone in the network.  
```{r eval=TRUE, echo=TRUE, message=FALSE}
res.closeness <- closeness(net)
res <- setDT(data.frame(res.closeness), keep.rownames = TRUE)[]
colnames(res)[1] <- "MSISDN"
colnames(res)[2] <- "Closeness"
res$MSISDN <- as.numeric(res$MSISDN)

res.closeness[1:5]
```
## 
####(2) Degree Centrality - the number of links connected to a node.
```{r eval=TRUE, echo=TRUE}
res <- cbind(res, degree(net))
colnames(res)[ncol(res)] <- "Degree"

degree(net)[1:5]
```
## 
####(3) Betweenness Centrality - the extent to which a node lies on shortest paths between other nodes.  
```{r eval=TRUE, echo=TRUE}
res <- cbind(res, betweenness(net))
colnames(res)[ncol(res)] <- "Betweenness"

betweenness(net)[1:5]
```
## 
####(4) Eigenvector Centrality - influence of a node in a network.
```{r eval=TRUE, echo=TRUE}
res <- cbind(res, eigen_centrality(net)$vector)
colnames(res)[ncol(res)] <- "Eigenvector"

eigen_centrality(net)$vector[1:5]
```
Result overview
```{r eval=TRUE, echo=TRUE}
head(res)
```

## 
####(5) Density - total number of edges over total edges
```{r eval=TRUE, echo=TRUE}
density <- edge_density(net, loops=FALSE)
density
```
## 
####(6) Reciprocity - the ratio of edges which are dual-directed over total edges
```{r eval=TRUE, echo=TRUE}
reciprocity <- reciprocity(net)
reciprocity
```

## Plots the network
```{r eval=TRUE, echo=TRUE}
V(net)$color=V(net)$name 
V(net)$color=gsub("82457","red","orange")
plot(net, edge.arrow.size=.4,vertex.label=NA)
```